---
title: "3d. Enrichment Analysis - Fig 2c"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load abundance data}
library(phyloseq)
library(ggpubr)
library(data.table)
library(tidyverse)
library(glue)
library(ggtext)

# This contains all the functional data necessary for cross-sectional analysis.
ps = readRDS("../Reference Files/functional_phyloseq.rds")
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# Load COG Key
cog_key = read.csv('../Reference Files/COG_key.csv') %>% 
  select(Cat,desc) %>% rename(Test=Cat) %>% mutate(desc = str_to_title(desc)) %>% 
  unique
# Load sig COG categories
cog = readxl::read_xlsx('../Results/Supplementary Data/Enrichment Analysis/COG2.xlsx') %>% 
  filter(qval<0.05) %>% 
  mutate(Dataset = 'COG') %>% left_join(cog_key) %>% 
  mutate(Test = desc)

# Load sig KEGG categories
ko = readxl::read_xlsx('../Results/Supplementary Data/Enrichment Analysis/KO2.xlsx',sheet='KEGG') %>% 
  filter(qval<0.05) %>% 
  mutate(Dataset = 'KEGG') %>% 
  separate(col = Test, into = c('L1','L2','L3','L4'), sep = ' [|] ')

# Load EC key
ec_key = read.csv('../Reference Files/EC_key_classes.csv')
# Load sig EC categories
ec = readxl::read_xlsx('../Results/Supplementary Data/Enrichment Analysis/EC2.xlsx') %>% 
  filter(qval<0.05) %>% 
  mutate(Dataset = 'EC')

# Combine datasets and format ~~~~~~~~~~~~~
enrich = cog %>% full_join(ko) %>% full_join(ec) %>% 
  mutate(Enrichment = log2(pct_pwy/pct_total),
         n_pwy = pwy_sig + pwy_ns) %>% 
  select(Dir,Level,Test,Dataset,L1,L2,L3,L4,Enrichment,n_pwy) %>% 
  mutate(Test = ifelse(Dataset=='KEGG' & !is.na(L4),L4,
                       ifelse(Dataset=='KEGG' & !is.na(L3),L3,
                              ifelse(Dataset=='KEGG' & !is.na(L2),L2,Test)))) %>% 
  arrange(Level) %>% arrange(L2) %>% arrange(Dataset) %>% 
  mutate(Enrichment = ifelse(Dir == 'Negative',Enrichment*-1,Enrichment))
```

```{r Formatting the table}
# Mainly formatting the names to make it as readable as possible.
enrich2 = enrich %>% 
  mutate(cat = ifelse(Dataset=='COG','Category',NA)) %>% 
  mutate(cat = ifelse(Dataset=='EC' & str_sub(Test,end=1)=='1',
                      ec_key$class.name[ec_key$Class=='1'] %>% unique,cat)) %>% 
  mutate(cat = ifelse(Dataset=='EC' & str_sub(Test,end=1)=='2',
                      ec_key$class.name[ec_key$Class=='2'] %>% unique,cat)) %>% 
  mutate(cat = ifelse(Dataset=='EC' & str_sub(Test,end=1)=='3',
                      ec_key$class.name[ec_key$Class=='3'] %>% unique,cat)) %>% 
  mutate(cat = ifelse(Dataset=='EC' & str_sub(Test,end=1)=='6',
                      ec_key$class.name[ec_key$Class=='6'] %>% unique,cat)) %>% 
  mutate(cat = ifelse(Dataset=='KEGG' & L2 == '09120 Genetic Information Processing',
                      'Gen. Info Processing',cat)) %>% 
  mutate(cat = ifelse(Dataset=='KEGG' & L2 == '09180 Brite Hierarchies',
                      'Brite Hierarchies',cat)) %>% 
  arrange(Test)

# Identify whether or not the test of interest is a child of another sig test
enrich2$prefix=''
for(i in 1:nrow(enrich2)){
  # i = 1
  t = enrich2$Level[i]
  if(t!='L1'){ # L1 can't be a child
    if(enrich2$Dataset[i]=='KEGG'){
      higher_lvl_n = paste0('L',as.character(as.numeric(str_sub(t,start=2))-1))
      higher_lvl_val = enrich2[[higher_lvl_n]] %>% .[i]
      while ((!is.null(higher_lvl_val)) & (higher_lvl_val %in% enrich2$Test)){
        enrich2$prefix[i] = paste0(enrich2$prefix[i],'\u2192 ')
        higher_lvl_n = paste0('L',as.character(as.numeric(str_sub(higher_lvl_n,start=2))-1))
        higher_lvl_val = enrich2[[higher_lvl_n]] %>% .[i]
      }
    } else if(enrich2$Dataset[i]=='EC'){
      split = str_split(enrich2$Test[i],'[.]')[[1]]
      higher_lvl_val = split[1:length(split)-1] %>% paste(collapse = '.')
      while ((!is.null(higher_lvl_val)) & (higher_lvl_val %in% enrich2$Test)){
        enrich2$prefix[i] = paste0(enrich2$prefix[i],'\u2192 ')
        split = str_split(higher_lvl_val,'[.]')[[1]]
        higher_lvl_val = split[1:length(split)-1] %>% paste(collapse = '.')
      }
    }
  }
  
}

# Transporters is a child of Brite Hierarchies, it just isn't a direct descendant so the above code doesn't work
enrich2$prefix[enrich2$Test=='02000 Transporters'] = '\u2192 '

enrich2 = enrich2 %>% 
                # Oxidoreductases
  mutate(Test = ifelse(Test=='1','1 Oxidoreductases',
                ifelse(Test=='1.1','1.1 Acting on CH-OH',
                ifelse(Test=='1.1.1','1.1.1 NAD(+)/NADP(+) acceptor',
                ifelse(Test=='1.8','1.8 Sulfur donor',
                # Transferases
                ifelse(Test=='2','2 Transferases',
                ifelse(Test=='2.4.2','2.4.2 Pentosyltransferases',
                ifelse(Test=='2.5','2.5 Non-methyl alkyl/aryl groups',
                ifelse(Test=='2.5.1','2.5.1 Non-methyl alkyl/aryl groups',
                ifelse(Test=='2.7','2.7 Phosphorus-containing groups',
                ifelse(Test=='2.7.2','2.7.2 Carboxy group acceptor',
                ifelse(Test=='2.7.4','2.7.4 Phosphate group acceptor',
                ifelse(Test=='2.7.7','2.7.7 Nucleotidyltransferases',
                ifelse(Test=='2.8','2.8 S-containing groups',
                ifelse(Test=='2.8.1','2.8.1 Sulfurtransferases',
                # Hydrolases
                ifelse(Test=='3.5.4','3.5.4 Acting on cyclic amidine C-N',
                # Ligases
                ifelse(Test=='6','6 Ligases',
                ifelse(Test=='6.1','6.1 Forming C-O bonds',
                ifelse(Test=='6.1.1','6.1.1 tRNA & related compounds',
                ifelse(Test=='6.3','6.3 Forming C-N bonds',
                ifelse(Test=='6.3.4','6.3.4 Other C-N ligases',
                # KEGG
                ifelse(Test=='09183 Protein families: signaling and cellular processes','09183 Signaling/Cellular Processes',
                ifelse(Test=='09182 Protein families: genetic information processing','09182 Gen. Info Processing',
                ifelse(Test=='09120 Genetic Information Processing','09120 Gen. Info Processing',
                ifelse(Test=='09130 Environmental Information Processing','09130 Env. Info Processing',Test))))))))))))))))))))))))) %>% 
  mutate(bold = ifelse(Test %in% c('1 Oxidoreductases',
                                   '2 Transferases',
                                   # Because it's the only one
                                   '3.5.4 Acting on cyclic amidine C-N',
                                   '6 Ligases',
                                   # KEGG
                                   '09180 Brite Hierarchies',
                                   '09120 Gen. Info Processing',
                                   # COG
                                   'Carbohydrate Transport And Metabolism'),
                       glue("<b>{Test}</b>"),Test)) %>% 
  arrange(rev(L4)) %>% arrange(rev(L3)) %>% arrange(rev(L2))

order = c(rev(enrich2$bold[enrich2$Dataset=='COG']),
          rev(enrich2$bold[enrich2$cat=='Oxidoreductases'] %>% sort()),
          rev(enrich2$bold[enrich2$cat=='Transferases'] %>% sort()),
          rev(enrich2$bold[enrich2$cat=='Hydrolases'] %>% sort()),
          rev(enrich2$bold[enrich2$cat=='Ligases'] %>% sort()),
          enrich2 %>% filter(Dataset=='KEGG',L2=='09180 Brite Hierarchies') %>% 
                arrange(L4) %>% arrange(L3) %>% pull(bold),
          enrich2 %>% filter(Dataset=='KEGG',L2=='09120 Genetic Information Processing') %>% 
                arrange(L4) %>% arrange(L3) %>% pull(bold))
enrich2$bold = factor(enrich2$bold, levels = order)

enrich2$cat = factor(enrich2$cat, levels = c('Category','Oxidoreductases','Transferases','Hydrolases','Ligases',sort(enrich2$cat[enrich2$Dataset=='KEGG'] %>% unique)))

enrich3 = enrich2 %>% 
  arrange(bold) %>%
  mutate(bold2 = as.character(bold)) %>% 
  mutate(bold2 = ifelse(!is.na(prefix) & str_sub(bold2,end=3) !='<b>',
                        paste0(prefix,bold2),bold2)) %>% 
  mutate(bold2 = ifelse(Test %in% c('1.1.1 NAD(+)/NADP(+) Acceptor'),paste0(prefix,bold2),bold2)) %>% 
  mutate(bolditalics = ifelse(prefix=='→ → ',paste0('<i>',bold2,'</i>'),bold2)) %>% 
  arrange(bold)

enrich3$bolditalics = factor(enrich3$bolditalics, levels = enrich3$bolditalics)
```

```{r}
library(ggh4x)
p = enrich3 %>%
  ggplot(aes(Enrichment,bolditalics,fill=Dir)) +
  geom_col() +
  theme_classic(base_size=16) +
  facet_nested(Dataset + cat ~ .,scales = 'free',space = 'free',
               strip = ggh4x::strip_nested(size = "variable")) +
  ylab(NULL) + xlab('Log 2 Fold Enrichment (Pathway/Background)') +
  theme(strip.text.y.right = element_text(angle = 0)) +
  theme(axis.text.y=element_markdown()) +
  theme(legend.position = 'none') +
  # xlim(-3.5,3.5) +
  theme(panel.spacing=unit(1,"lines"))

ggsave('../Results/2. Differential Abundance & Enrichments/Fig. 2c - Enriched and Depleted Functions.jpeg',plot = p,height=12, width = 10)

plotdata = p$data

writexl::write_xlsx(list(plotdata),"C:/Users/armetcal/OneDrive - UBC/Grad School/Data/HUMAN COHORT/THESIS/Tables/CHAP 3/Table A.3.5 - Enrichment Data.xlsx")
```
