---
title: "6b. Disease Progression - Species"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load data}
library(phyloseq)
library(tidyverse)
library(ggpubr)
library(microbiome)
library(ggcorrplot)
library(lme4)
library(lmerTest)
library(scales)
```

# LONGITUDINAL ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# SPECIES

```{r}
# Prepare the data ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# We will combine our longitudinal dataset with our taxonomy data, where only the microbes of interest are included.

# Variables to test
vars = c('mds1.total','mds2.total','mds3.total','mds4.total','levo.eq.dose')

# Load taxonomic data, CLR transform
ps = readRDS('../Reference Files/taxonomy_phyloseq.rds') %>% microbiome::transform('clr')

# Load sig taxa
# Multivar
temp2 = readxl::read_xlsx('../Results/Supplementary Data/Differential Abundance/Taxonomy_Multivar_Summary.xlsx',sheet='Species') %>%
  filter(sig_in_at_least_n==1, Variable=='StatusPD')

# Make a vector of the species to test
# Include F. prausnitzii because of its strong association with 
#  most Ctrl-associated pathways (Table 2)
sig = c(temp2[['Species']],'s__Faecalibacterium_prausnitzii') %>% unique
# sig = c(temp2[['Species']],'s__Faecalibacterium_prausnitzii')

# Select species and melt phyloseq object
other_levels = colnames(ps@tax_table)[colnames(ps@tax_table) != 'Species']
psm = ps %>% subset_taxa(Species %in% sig) %>% psmelt() %>%
  select(Sample,Status,Sex,laxatives,depth,OTU,Abundance,Kingdom:Species) %>% unique
# Load longitudinal metadata
long0 = read.csv('../Reference Files/longitudinal_data.csv') %>% 
  select(Sample,redcap_event_name,months_last_visit,all_of(vars))

# Finished longitudinal dataset. Contains all the visits for each person with n>1 visits,
# where each individual visit/microbe combination is one row.
long = long0 %>% left_join(psm, multiple='all')

# Center static predictors, normalize where necessary.
# This reduces spurious correlations as explained in LMER documentation.
# Center each OTU separately.
pd = long %>% filter(Status == 'PD') %>% group_by(OTU) %>% 
  # Rescale to range of 0-1
  mutate(Abundance = rescale(Abundance)) %>% 
  # Center around 0
  mutate(Abundance = Abundance-mean(Abundance,na.rm=T)) %>% ungroup()

# Rescale to 0-1
pd = pd %>% mutate_at(vars('months_last_visit',all_of(vars),'depth',contains('base.')),rescale)
# 
# # Center scale
pd = pd %>% mutate_at(vars('months_last_visit',all_of(vars),'depth',contains('base.')),
                      function(x) x - mean(x,na.rm=T))
```

## 'Univariate' model that only controls for age and baseline

```{r}
# Univar
stats.uni = tibble()
for(v in vars){
  for(s in sig){
    # s = sig[1]
    temp = pd %>% filter(Species == s)
    names(temp)[names(temp)==v] = 'test'
    
    lmertest = lmer(test ~ months_last_visit*Abundance + (months_last_visit|Sample), data = temp,
                    REML = F,
                    control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
    
    L = lmertest %>% broom.mixed::tidy() %>% as.tibble() %>% mutate(Variable = v, Species = s)
    stats.uni = stats.uni %>% rbind(L)
  }
}

stats.uni.filt = stats.uni %>% filter(str_detect(term,'months_last_visit:Abundance')) %>%
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup
```

## 'Multivar' model that controls for relevant microbial and non-microbial variables

```{r}
# Univar
stats.multi = tibble()
for(v in vars){
  for(s in sig){
    # s = sig[1]
    temp = pd %>% filter(Species == s)
    names(temp)[names(temp)==v] = 'test'
    
    lmertest = lmer(test ~ months_last_visit*Abundance + laxatives + Sex + depth + (months_last_visit|Sample), data = temp,
                    REML = F,
                    control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
    
    L = lmertest %>% broom.mixed::tidy() %>% as.tibble() %>% mutate(Variable = v, Species = s)
    stats.multi = stats.multi %>% rbind(L)
  }
}

stats.multi.filt = stats.multi %>% filter(str_detect(term,'months_last_visit:Abundance')) %>% 
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup
```


## Plot

```{r}
tempstats.multi = stats.multi %>% filter(term=='months_last_visit:Abundance') %>% 
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup %>% 
  mutate(Qval = ifelse(q.value>0.05,'',
                       ifelse(q.value>0.01,'*',
                              ifelse(q.value>0.001,'**',
                                     ifelse(q.value<=0.001,'***',''))))) %>% 
  mutate(Qval2 = ifelse(q.value>0.1,'',
                        ifelse(q.value>0.05,'+',''))) %>% 
  rename('Estimate' = estimate)

# Edit levels for plotting
tempstats.multi$Variable[tempstats.multi$Variable=='levo.eq.dose'] = 'Levodopa\nEq. Dose'
tempstats.multi$Variable[tempstats.multi$Variable=='mds1.total'] = 'UPDRS 1'
tempstats.multi$Variable[tempstats.multi$Variable=='mds2.total'] = 'UPDRS 2'
tempstats.multi$Variable[tempstats.multi$Variable=='mds3.total'] = 'UPDRS 3'
tempstats.multi$Variable[tempstats.multi$Variable=='mds4.total'] = 'UPDRS 4'
tempstats.multi$Species = str_remove_all(tempstats.multi$Species,'s__') %>% str_replace_all('_',' ')

# Arrange data
order_species = tempstats.multi %>% filter(Variable=='Levodopa\nEq. Dose') %>% arrange(`Estimate`)
tempstats.multi$Species = factor(tempstats.multi$Species, levels = order_species$Species)

tempstats.multi$Variable = factor(tempstats.multi$Variable,levels = c('Levodopa\nEq. Dose','UPDRS 1','UPDRS 2','UPDRS 3','UPDRS 4'))
tempstats.multi = tempstats.multi %>% droplevels() %>% as.data.frame() %>% filter(!is.na(Variable))

plot_a = tempstats.multi %>% ggplot(aes(Variable,Species,col=`Estimate`)) +
  geom_point(size=14) +
  theme_classic(base_size=18) +
  ylab(NULL) + xlab(NULL) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
  geom_text(aes(label=Qval),size = 12, col = 'white',nudge_y = -0.25) +
  geom_text(aes(label=Qval2),size = 12, col = 'white',nudge_y = 0)
```

# METACYC

```{r}
# MetaCyc Data - CLR transformed
ps = readRDS("../Reference Files/functional_phyloseq.rds") %>% .[["pwy"]] %>% microbiome::transform('clr')

# Load sig pwy
temp2 = readxl::read_xlsx('../Results/Supplementary Data/Differential Abundance/MetaCyc_Summary.xlsx',sheet='multivar') %>%
  filter(sig_in_at_least_n==1, Variable=='StatusPD')

# Make a vector of the pathways to test
sig = temp2[['Species']]

# Select species and melt phyloseq object
other_levels = colnames(ps@tax_table)[colnames(ps@tax_table) != 'Species']
psm = ps %>% subset_taxa(Species %in% sig) %>% psmelt() %>%
  select(Sample,Status,Sex,laxatives,depth,OTU,Abundance,Species) %>% unique
# Load longitudinal metadata
long0 = read.csv('../Reference Files/longitudinal_data.csv') %>% 
  select(-Disease.duration,-Status)
table(long0$redcap_event_name)

# Finished longitudinal dataset. Contains all the visits for each person with n>1 visits,
# where each individual visit/microbe combination is one row.
# baseline bristol rating and microbial abundances are static variables.
long = long0 %>% left_join(psm, multiple='all')
attach(long %>% select(Status,redcap_event_name,Sample) %>% unique)
table(Status,redcap_event_name)
detach()

# Center static predictors, normalize where necessary.
# This reduces spurious correlations as explained in LMER documentation.
# Center each OTU separately.
pd = long %>% filter(Status == 'PD') %>% group_by(OTU) %>% 
  # Rescale to range of 0-1
  mutate(Abundance = rescale(Abundance)) %>% 
  # Center around 0
  mutate(Abundance = Abundance-mean(Abundance,na.rm=T)) %>% ungroup()

# Rescale to 0-1
pd = pd %>% mutate_at(vars('months_last_visit',all_of(vars),'depth',contains('base.')),rescale)
# 
# # Center scale
pd = pd %>% mutate_at(vars('months_last_visit',all_of(vars),'depth',contains('base.')),
                      function(x) x - mean(x,na.rm=T))
```

## 'Univariate' model that only controls for age and baseline

```{r}
# Univar
stats.uni = tibble()
for(v in vars){
  for(s in sig){
    # v = vars[1]
    # s = sig[1]
    temp = pd %>% filter(Species == s)
    names(temp)[names(temp)==v] = 'test'
    
    lmertest = lmer(test ~ months_last_visit*Abundance + (months_last_visit|Sample), data = temp,
                    REML = F,
                    control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
    
    L = lmertest %>% broom.mixed::tidy() %>% as.tibble() %>% mutate(Variable = v, Species = s)
    stats.uni = stats.uni %>% rbind(L)
  }
}

stats.uni.filt = stats.uni %>% filter(str_detect(term,'months_last_visit:Abundance')) %>%
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup
```

## 'Multivar' model that controls for relevant microbial and non-microbial variables

```{r}
stats.multi = tibble()
for(v in vars){
  for(s in sig){
    # s = sig[1]
    temp = pd %>% filter(Species == s)
    names(temp)[names(temp)==v] = 'test'
    
    lmertest = lmer(test ~ months_last_visit*Abundance + laxatives + Sex + depth + (months_last_visit|Sample), data = temp,
                    REML = F,
                    control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000)))
    
    L = lmertest %>% broom.mixed::tidy() %>% as.tibble() %>% mutate(Variable = v, Species = s)
    stats.multi = stats.multi %>% rbind(L)
  }
}

stats.multi.filt = stats.multi %>% filter(str_detect(term,'months_last_visit:Abundance')) %>% 
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup
```

## Plot

```{r}
tempstats.multi = stats.multi %>% filter(term=='months_last_visit:Abundance') %>% 
  group_by(Variable,term) %>% 
  mutate(q.value = p.adjust(p.value, method='BH')) %>% ungroup %>% 
  mutate(Qval = ifelse(q.value>0.05,'',
                       ifelse(q.value>0.01,'*',
                              ifelse(q.value>0.001,'**',
                                     ifelse(q.value<=0.001,'***',''))))) %>% 
  mutate(Qval2 = ifelse(q.value>0.1,'',
                        ifelse(q.value>0.05,'+',''))) %>% 
  rename('Estimate' = estimate) %>%
  mutate(Species = str_replace(Species, '[&]beta;-D','B-D')) %>% 
  mutate(Species = str_wrap(Species,40))

# Edit levels for plotting
tempstats.multi$Variable[tempstats.multi$Variable=='levo.eq.dose'] = 'Levodopa\nEq. Dose'
tempstats.multi$Variable[tempstats.multi$Variable=='mds1.total'] = 'UPDRS 1'
tempstats.multi$Variable[tempstats.multi$Variable=='mds2.total'] = 'UPDRS 2'
tempstats.multi$Variable[tempstats.multi$Variable=='mds3.total'] = 'UPDRS 3'
tempstats.multi$Variable[tempstats.multi$Variable=='mds4.total'] = 'UPDRS 4'
tempstats.multi$Species = str_remove_all(tempstats.multi$Species,'s__') %>% str_replace_all('_',' ')

tempstats.multi$Variable = factor(tempstats.multi$Variable,levels = c('Levodopa\nEq. Dose','UPDRS 1','UPDRS 2','UPDRS 3','UPDRS 4'))
tempstats.multi = tempstats.multi %>% droplevels() %>% as.data.frame() %>% filter(!is.na(Variable))
  
# Arrange data
order_meta = tempstats.multi %>% filter(Variable=='Levodopa\nEq. Dose') %>% arrange(`Estimate`)
tempstats.multi$Species = factor(tempstats.multi$Species, levels = order_meta$Species)

plot_b = tempstats.multi %>% ggplot(aes(Variable,Species,col=`Estimate`)) +
  geom_point(size=14) +
  theme_classic(base_size=18) +
  ylab(NULL) + xlab(NULL) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
  geom_text(aes(label=Qval),size = 12, col = 'white',nudge_y = -0.25) +
  geom_text(aes(label=Qval2),size = 12, col = 'white',nudge_y = 0)
```

# BASELINE SPECIES ~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# Taxonomy Data - CLR transformed
ps = readRDS("../Reference Files/taxonomy_phyloseq.rds") %>% microbiome::transform('clr')
# Load functions that allow high-throughput Spearman & GLM analysis
source('999. Cross-Sectional Correlation Functions.R')

# Load sig taxa
temp2 = readxl::read_xlsx('../Results/Supplementary Data/Differential Abundance/Taxonomy_Multivar_Summary.xlsx',sheet='Species') %>%
  filter(sig_in_at_least_n==1, Variable=='StatusPD')

# Make a vector of the species to test
# Include F. prausnitzii because of its strong association with 
#  Ctrl-associated pathways (Table 2)
sig = c(temp2[['Species']],'s__Faecalibacterium_prausnitzii')

# Select species and melt phyloseq object
other_levels = colnames(ps@tax_table)[colnames(ps@tax_table) != 'Species']
psm = ps %>% subset_taxa(Species %in% sig) %>% psmelt() %>%
  select(Species,everything()) %>% select(-any_of(other_levels),-OTU) %>%
  pivot_wider(names_from = Species, values_from = Abundance) %>% 
  select(-any_of(c('bdi_total','fss_total','mds1.total','mds2.total','mds3.total','mds4.total','levo.eq.dose')))

# Load longitudinal metadata
long0 = read.csv('../Reference Files/longitudinal_data.csv') %>% 
  select(-Disease.duration,-Status)
long = long0 %>% filter(months_last_visit==0) %>% unique()

# Combine with CRP and metabolomic data
psm = psm %>% left_join(long)

# Filter for only PD
psm = psm %>% filter(Status=='PD')
```

## Univariate Analysis

```{r, include=FALSE}
spearman = correlate_met(psm %>% filter(!is.na(mds1.total),!is.na(Disease.duration)),
                         'mds1.total',grp=NA) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds2.total),!is.na(Disease.duration)),
                          'mds2.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds3.total),!is.na(Disease.duration)),
                          'mds3.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds4.total),!is.na(Disease.duration)),
                           'mds4.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(bdi_total),!is.na(Disease.duration)),
                           'bdi_total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(fss_total),!is.na(Disease.duration)),
                           'fss_total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(levo.eq.dose),!is.na(Disease.duration)),
                           'levo.eq.dose',grp=NA))
```

## Multivariate Analysis

```{r}
glm = glm_met_severity(psm %>% filter(!is.na(mds1.total),!is.na(Disease.duration)),
                           'mds1.total',grp=NA) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds2.total),!is.na(Disease.duration)),
                           'mds2.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds3.total),!is.na(Disease.duration)),
                           'mds3.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds4.total),!is.na(Disease.duration)),
                           'mds4.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(bdi_total),!is.na(Disease.duration)),
                           'bdi_total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(fss_total),!is.na(Disease.duration)),
                           'fss_total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(levo.eq.dose),!is.na(Disease.duration)),
                           'levo.eq.dose',grp=NA))

# Extract just the relevant rows for plotting
glm_species = glm %>% filter(str_detect(Explanatory,'Microbe'))
```

## Multivar Plots

```{r}
# Edit levels for plottin
glm_species$Response[glm_species$Response=='mds1.total'] = 'UPDRS 1'
glm_species$Response[glm_species$Response=='mds2.total'] = 'UPDRS 2'
glm_species$Response[glm_species$Response=='mds3.total'] = 'UPDRS 3'
glm_species$Response[glm_species$Response=='mds4.total'] = 'UPDRS 4'
glm_species$Response[glm_species$Response=='levo.eq.dose'] = 'Levodopa\nEq. Dose'
glm_species$Response[glm_species$Response=='bdi_total'] = 'Total BDI'
glm_species$Response[glm_species$Response=='fss_total'] = 'Total FSS'
glm_species$Taxon = str_remove_all(glm_species$Taxon,'s__') %>% str_replace_all('_',' ')

glm_species$Response = factor(glm_species$Response,levels =  c('Levodopa\nEq. Dose','UPDRS 1','UPDRS 2','UPDRS 3','UPDRS 4','Total BDI','Total FSS'))

# Plot ordering
order = glm_species %>% filter(Response=='Levodopa\nEq. Dose') %>% 
  arrange(Estimate)
order$Taxon = str_remove_all(order$Taxon,'s__') %>% str_replace_all('_',' ')
glm_species$Taxon = factor(glm_species$Taxon, levels = order_species$Species)

# Plotting function - labels the plot automatically
# vars = the variables you want to include in the plot
prep_plot = function(vars){
  p = glm_species %>% 
    mutate(Qval = ifelse(qval>0.05,'',
                        ifelse(qval>0.01,'*',
                               ifelse(qval>0.001,'**',
                                      ifelse(qval<=0.001,'***',''))))) %>% 
    mutate(Qval2 = ifelse(qval>0.1,'',
                         ifelse(qval>0.05,'+',''))) %>% 
    filter(Response %in% vars) %>% 
    ggplot(aes(Response,Taxon,col = Estimate)) +
    geom_point(size =14) +
    theme_classic(base_size = 16) +
    scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
    geom_text(aes(label=Qval),size = 10, col = 'white',nudge_y = -0.25) +
    geom_text(aes(label=Qval2),size = 10, col = 'white',nudge_y = 0) +
    ylab(NULL) + xlab(NULL) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}

# Plots
plot_c = prep_plot(unique(glm_species$Response))
```

# BASELINE METACYC ~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
# MetaCyc Data
ps = readRDS("../Reference Files/functional_phyloseq.rds") %>% .[["pwy"]] %>% microbiome::transform('clr')
# Load functions that allow high-throughput Spearman & GLM analysis
source('999. Cross-Sectional Correlation Functions.R')
```

```{r}
# Load sig pwy
temp2 = readxl::read_xlsx('../Results/Supplementary Data/Differential Abundance/MetaCyc_Summary.xlsx',sheet='multivar') %>%
  filter(sig_in_at_least_n==1, Variable=='StatusPD')

# Make a vector of the pathways to test
sig = temp2[['Species']]

# Select species and melt phyloseq object
other_levels = colnames(ps@tax_table)[colnames(ps@tax_table) != 'Species']
psm = ps %>% subset_taxa(Species %in% sig) %>% psmelt() %>%
  select(Species,everything()) %>% select(-any_of(other_levels),-OTU) %>%
  pivot_wider(names_from = Species, values_from = Abundance) %>% 
  select(-any_of(c('bdi_total','fss_total','mds1.total','mds2.total','mds3.total','mds4.total','levo.eq.dose')))

# Load longitudinal metadata
long0 = read.csv('../Reference Files/longitudinal_data.csv') %>% 
  select(-Disease.duration,-Status)
long = long0 %>% filter(months_last_visit==0) %>% unique()

# Combine with CRP and metabolomic data
psm = psm %>% left_join(long)

# Filter for only PD
psm = psm %>% filter(Status=='PD')
```

## Univariate Analysis

```{r}
spearman = correlate_met(psm %>% filter(!is.na(mds1.total),!is.na(Disease.duration)),
                         'mds1.total',grp=NA) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds2.total),!is.na(Disease.duration)),
                          'mds2.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds3.total),!is.na(Disease.duration)),
                          'mds3.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(mds4.total),!is.na(Disease.duration)),
                           'mds4.total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(bdi_total),!is.na(Disease.duration)),
                           'bdi_total',grp=NA)) %>% 
  rbind(correlate_met(psm %>% filter(!is.na(fss_total),!is.na(Disease.duration)),
                           'fss_total',grp=NA)) %>%
  rbind(correlate_met(psm %>% filter(!is.na(levo.eq.dose),!is.na(Disease.duration)),
                           'levo.eq.dose',grp=NA)) %>%
  mutate(taxon = str_replace(taxon, '[&]beta;-D','B-D')) %>% 
  mutate(taxon = str_wrap(taxon,40))
```

## Multivariate Analysis

```{r}
glm = glm_met_severity(psm %>% filter(!is.na(mds1.total),!is.na(Disease.duration)),
                           'mds1.total',grp=NA) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds2.total),!is.na(Disease.duration)),
                           'mds2.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds3.total),!is.na(Disease.duration)),
                           'mds3.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(mds4.total),!is.na(Disease.duration)),
                           'mds4.total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(bdi_total),!is.na(Disease.duration)),
                           'bdi_total',grp=NA)) %>% 
  rbind(glm_met_severity(psm %>% filter(!is.na(fss_total),!is.na(Disease.duration)),
                           'fss_total',grp=NA)) %>%
  rbind(glm_met_severity(psm %>% filter(!is.na(levo.eq.dose),!is.na(Disease.duration)),
                           'levo.eq.dose',grp=NA)) %>%
  mutate(Taxon = str_replace(Taxon, '[&]beta;-D','B-D')) %>% 
  mutate(Taxon = str_wrap(Taxon,40))

# Extract just the relevant rows for plotting
glm_species = glm %>% filter(str_detect(Explanatory,'Microbe'))
```

## Multivar Plots

```{r}
# Edit levels for plotting
glm_species$Response[glm_species$Response=='mds1.total'] = 'UPDRS 1'
glm_species$Response[glm_species$Response=='mds2.total'] = 'UPDRS 2'
glm_species$Response[glm_species$Response=='mds3.total'] = 'UPDRS 3'
glm_species$Response[glm_species$Response=='mds4.total'] = 'UPDRS 4'
glm_species$Response[glm_species$Response=='levo.eq.dose'] = 'Levodopa\nEq. Dose'
glm_species$Response[glm_species$Response=='bdi_total'] = 'Total BDI'
glm_species$Response[glm_species$Response=='fss_total'] = 'Total FSS'
glm_species$Taxon = str_remove_all(glm_species$Taxon,'s__') %>% str_replace_all('_',' ')

glm_species$Response = factor(glm_species$Response,levels = c('Levodopa\nEq. Dose','UPDRS 1','UPDRS 2','UPDRS 3','UPDRS 4','Total BDI','Total FSS'))

# Plotting function - labels the plot automatically
# vars = the variables you want to include in the plot
prep_plot = function(vars){
  p = glm_species %>% 
    mutate(Qval = ifelse(qval>0.05,'',
                        ifelse(qval>0.01,'*',
                               ifelse(qval>0.001,'**',
                                      ifelse(qval<=0.001,'***',''))))) %>% 
    mutate(Qval2 = ifelse(qval>0.1,'',
                         ifelse(qval>0.05,'+',''))) %>% 
    filter(Response %in% vars) %>% 
    ggplot(aes(Response,Taxon,col = Estimate)) +
    geom_point(size =14) +
    theme_classic(base_size = 16) +
    scale_color_gradient2(low="darkblue", high="darkred", guide="colorbar") +
    geom_text(aes(label=Qval),size = 10, col = 'white',nudge_y = -0.25) +
    geom_text(aes(label=Qval2),size = 10, col = 'white',nudge_y = 0) +
    ylab(NULL) + xlab(NULL) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  return(p)
}

glm_species$Taxon = factor(glm_species$Taxon, levels = order_meta$Species)

# Plots
plot_d = prep_plot(unique(glm_species$Response))
```

# COMPILE PLOT

```{r}
p1 = ggarrange(plotlist = list(NULL,plot_a, NULL,plot_c), ncol = 4, widths = c(0.7,3.5,0.7,4), common.legend = F)
p1
p2 = ggarrange(plotlist = list(plot_b, plot_d), ncol = 2, widths = c(3.8,4), common.legend = F)
p2
p3 = ggarrange(plotlist = list(p1, p2), ncol = 1, nrow = 2, heights = c(1,1.1))
p3

ggsave('../Results/Figures/Figure 4 - Progression.pdf',height = 10.5, width = 18)
```

