---
title: "Figure 5 - Random Forest"
author: "Avril Metcalfe-Roach"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(phyloseq)
library(ggpubr)
library(data.table)
library(tidyverse)
library(readxl)
library(scales)
```

# Importance Values - Status

```{r Formatting function and dataframes}
format_features = function(v){
  temp = sapply(v,function(x){
    if(str_detect(x,'s__')){
      a = x %>% str_remove('s__') %>% str_split('_') %>% .[[1]]
      a[1] = paste(str_sub(a[1],end=1),'.')
      return(paste(a[1],a[2]))
    } else {
      return(str_wrap(x,20))
    }
  })
  return(temp)
}
fix = tibble(Feature = c('Sex','smokeprev_or_before_pd','bristol','coffee_per_day'),
             fix = c('Sex (Male)', "Ever Smoked",'Bristol Rating','Coffee/Day'))
meta_names = tibble(Feature = c("PWY-7383: anaerobic energy metabolism (invertebrates, cytosol)",
                                "PWY-6471: peptidoglycan biosynthesis IV (Enterococcus faecium)",
                                "PWY66-409: superpathway of purine nucleotide salvage",
                                "PWY-5177: glutaryl-CoA degradation",     
                                "PWY-7242: D-fructuronate degradation",     
                                "GLUCUROCAT-PWY: superpathway of &beta;-D-glucuronide and D-glucuronate degradation",
                                "GALACT-GLUCUROCAT-PWY: superpathway of hexuronide and hexuronate degradation"),
                    new = c("PWY-7383:\nanaerobic metabolism",
                                "PWY-6471:\npeptidoglycan biosynth.",
                                "PWY66-409:\npurine salvage",
                                "PWY-5177:\nglutaryl-CoA degrad.",     
                                "PWY-7242:\nfructuronate degrad.",     
                                "GLUCUROCAT-PWY:\nglucuronide/ate degrad.",
                                "GALACT-GLUCUROCAT-PWY:\nhexuronide/ate degrad."))
key = read.csv('../Reference Files/EC_key.csv') %>% rename(Feature=EC) %>% select(-EC2) %>% 
  mutate(ec = sapply(ec, function(x) x %>% str_split(' // ') %>% .[[1]] %>% .[1]))
```

```{r Formatted Importance Values}
# Load Importance Values
imp = read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - Covars.xlsx') %>% mutate(Data='Covar') %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - Taxonomy.xlsx', 
                      sheet='Taxonomy') %>% mutate(Data='Taxonomy')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - Taxonomy.xlsx', 
                      sheet='Taxonomy & Covar') %>% mutate(Data='Taxonomy & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - MetaCyc.xlsx', 
                      sheet='MetaCyc') %>% mutate(Data='MetaCyc')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - MetaCyc.xlsx', 
                      sheet='MetaCyc & Covar') %>% mutate(Data='MetaCyc & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - KO.xlsx', 
                      sheet='KO') %>% mutate(Data='KO')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - KO.xlsx', 
                      sheet='KO & Covar') %>% mutate(Data='KO & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - COG.xlsx', 
                      sheet='COG') %>% mutate(Data='COG')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - COG.xlsx', 
                      sheet='COG & Covar') %>% mutate(Data='COG & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - EC.xlsx', 
                      sheet='EC') %>% mutate(Data='EC')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - EC.xlsx', 
                      sheet='EC & Covar') %>% mutate(Data='EC & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - Metabolites Annot&Covar.xlsx', 
                      sheet='Metabolites') %>% mutate(Data='Metabolites')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Status - Metabolites Annot&Covar.xlsx', 
                      sheet='Metabolites & Covar') %>% mutate(Data='Metabolites & Covar')) %>% 
  mutate(Feature = ifelse(Feature=='logcoffeeperday','coffee_per_day',Feature))

# Add direction info and format the Feature names for plotting
imp2 = imp %>%
  left_join(fix) %>% mutate(Feature = ifelse(!is.na(fix),fix,Feature)) %>% 
  left_join(meta_names) %>% mutate(Feature = ifelse(!is.na(new),new,Feature)) %>% 
  mutate(Feature = format_features(.$Feature)) 
imp2$Feature[str_detect(imp2$Data,'EC|COG|KO')] = sapply(imp2$Feature[str_detect(imp2$Data,'EC|COG|KO')],
                                                            function(x) x %>% str_split(':') %>% .[[1]] %>% .[1])
imp2 = imp2 %>% mutate(Feature = ifelse(str_detect(Data,'EC|COG|KO'), str_remove_all(Feature,' |\n'),Feature))
imp2 = imp2 %>% 
  left_join(key) %>% 
  mutate(Feature = ifelse(!is.na(ec),ec,Feature)) %>% 
  select(-ec,-new,-fix)

# Order by magnitude
imp3 = imp2 %>% 
  arrange(Importance) %>% 
  mutate(Data = factor(Data, levels = c('Covar',sort(unique(.$Data))[sort(unique(.$Data))!='Covar']))) %>% 
  arrange(Data) %>% 
  select(-`...1`)

# For reference below:
imp4 = imp2 %>% rename(fixed = Feature) %>% left_join(imp)
```

```{r Save Importance Table}
# L = list(Taxonomy = imp3 %>% filter(str_detect(Data,'Taxonomy') | Data=='Covar'),
#          MetaCyc = imp3 %>% filter(str_detect(Data,'MetaCyc') | Data=='Covar'),
#          KO = imp3 %>% filter(str_detect(Data,'KO') | Data=='Covar'),
#          COG = imp3 %>% filter(str_detect(Data,'COG') | Data=='Covar'),
#          EC = imp3 %>% filter(str_detect(Data,'EC') | Data=='Covar'),
#          Metabolites = imp3 %>% filter(str_detect(Data,'Metabolites') | Data=='Covar'))
# 
# writexl::write_xlsx(L,'../Results/Supplementary Data/Random Forest/00. RF Importance Values - Status - Formatted.xlsx')
```

```{r Plot Importance Values}
# Taxonomy
p1 = imp3 %>% filter(Data=='Covar') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p2 = imp3 %>% filter(Data=='Taxonomy') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Taxonomy') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p3 = imp3 %>% filter(Data=='Taxonomy & Covar') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Taxonomy & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

plot.taxa = ggarrange(p1,p2,p3, ncol=3, widths = c(4,6,8))

# MetaCyc
p4 = imp3 %>% filter(Data=='MetaCyc') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('MetaCyc') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='MetaCyc & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('MetaCyc & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

plot.meta = ggarrange(p4,p5, ncol=2, widths = c(6,8))

# KO
p4 = imp3 %>% filter(Data=='KO') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('KO') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='KO & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('KO & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

plot.ko = ggarrange(p4,p5, nrow=2)

# EC
p4 = imp3 %>% filter(Data=='EC') %>%
  .[-which(duplicated(.$Feature)),] %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('EC') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='EC & Covar') %>%
  .[-which(duplicated(.$Feature)),] %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('EC & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

plot.ec = ggarrange(p4,p5, nrow=2)

# COG
p4 = imp3 %>% filter(Data=='COG') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('COG') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='COG & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('COG & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

plot.cog = ggarrange(p4,p5, nrow=2)
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Importance Values - Progression

```{r Formatting function and dataframes}
format_features = function(v){
  temp = sapply(v,function(x){
    if(str_detect(x,'s__')){
      a = x %>% str_remove('s__') %>% str_split('_') %>% .[[1]]
      a[1] = paste(str_sub(a[1],end=1),'.')
      return(paste(a[1],a[2]))
    } else {
      return(str_wrap(x,20))
    }
  })
  return(temp)
}
fix = tibble(Feature = c('Sex','smokeprev_or_before_pd','bristol','coffee_per_day'),
             fix = c('Sex (Male)', "Ever Smoked",'Bristol Rating','Coffee/Day'))
meta_names = tibble(Feature = c("PWY-7383: anaerobic energy metabolism (invertebrates, cytosol)",
                                "PWY-6471: peptidoglycan biosynthesis IV (Enterococcus faecium)",
                                "PWY66-409: superpathway of purine nucleotide salvage",
                                "PWY-5177: glutaryl-CoA degradation",     
                                "PWY-7242: D-fructuronate degradation",     
                                "GLUCUROCAT-PWY: superpathway of &beta;-D-glucuronide and D-glucuronate degradation",
                                "GALACT-GLUCUROCAT-PWY: superpathway of hexuronide and hexuronate degradation"),
                    new = c("PWY-7383:\nanaerobic metabolism",
                                "PWY-6471:\npeptidoglycan biosynth.",
                                "PWY66-409:\npurine salvage",
                                "PWY-5177:\nglutaryl-CoA degrad.",     
                                "PWY-7242:\nfructuronate degrad.",     
                                "GLUCUROCAT-PWY:\nglucuronide/ate degrad.",
                                "GALACT-GLUCUROCAT-PWY:\nhexuronide/ate degrad."))
key = read.csv('../Reference Files/EC_key.csv') %>% rename(Feature=EC) %>% select(-EC2) %>% 
  mutate(ec = sapply(ec, function(x) x %>% str_split(' // ') %>% .[[1]] %>% .[1]))
```

```{r Formatted Importance Values}
# Load Importance Values
imp = read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - Covars.xlsx') %>% mutate(Data='Covar') %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - Taxonomy.xlsx', 
                      sheet='Taxonomy') %>% mutate(Data='Taxonomy')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - Taxonomy.xlsx', 
                      sheet='Taxonomy & Covar') %>% mutate(Data='Taxonomy & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - MetaCyc.xlsx', 
                      sheet='MetaCyc') %>% mutate(Data='MetaCyc')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - MetaCyc.xlsx', 
                      sheet='MetaCyc & Covar') %>% mutate(Data='MetaCyc & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - KO.xlsx', 
                      sheet='KO') %>% mutate(Data='KO')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - KO.xlsx', 
                      sheet='KO & Covar') %>% mutate(Data='KO & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - COG.xlsx', 
                      sheet='COG') %>% mutate(Data='COG')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - COG.xlsx', 
                      sheet='COG & Covar') %>% mutate(Data='COG & Covar')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - EC.xlsx', 
                      sheet='EC') %>% mutate(Data='EC')) %>% 
  full_join(read_xlsx('../Results/Supplementary Data/Random Forest/RF Importance Values - Progression - EC.xlsx', 
                      sheet='EC & Covar') %>% mutate(Data='EC & Covar')) %>% 
  mutate(Feature = ifelse(Feature=='logcoffeeperday','coffee_per_day',Feature))

# Add direction info and format the Feature names for plotting
imp2 = imp %>%
  left_join(fix) %>% mutate(Feature = ifelse(!is.na(fix),fix,Feature)) %>% 
  left_join(meta_names) %>% mutate(Feature = ifelse(!is.na(new),new,Feature)) %>% 
  mutate(Feature = format_features(.$Feature)) 
imp2$Feature[str_detect(imp2$Data,'EC|COG|KO')] = sapply(imp2$Feature[str_detect(imp2$Data,'EC|COG|KO')],
                                                            function(x) x %>% str_split(':') %>% .[[1]] %>% .[1])
imp2 = imp2 %>% mutate(Feature = ifelse(str_detect(Data,'EC|COG|KO'), str_remove_all(Feature,' |\n'),Feature))
imp2 = imp2 %>% mutate(Feature = ifelse(str_detect(Feature,'ARGININE-DEIMINASE'),'ARGININE-DEIMINASE-RXN',Feature)) %>% 
  left_join(key) %>% 
  mutate(Feature = ifelse(!is.na(ec),ec,Feature)) %>% 
  select(-ec,-new,-fix)

imp3 = imp2 %>%
  arrange(Importance) %>% 
  mutate(Data = factor(Data, levels = c('Covar',sort(unique(.$Data))[sort(unique(.$Data))!='Covar']))) %>% 
  arrange(Data) %>% 
  select(-`...1`)

# For reference below:
imp4 = imp2 %>% rename(fixed = Feature) %>% left_join(imp)
```

```{r Save Importance Table}
# L = list(Taxonomy = imp3 %>% filter(str_detect(Data,'Taxonomy') | Data=='Covar'),
#          MetaCyc = imp3 %>% filter(str_detect(Data,'MetaCyc') | Data=='Covar'),
#          KO = imp3 %>% filter(str_detect(Data,'KO') | Data=='Covar'),
#          COG = imp3 %>% filter(str_detect(Data,'COG') | Data=='Covar'),
#          EC = imp3 %>% filter(str_detect(Data,'EC') | Data=='Covar'))
# 
# writexl::write_xlsx(L,'../Results/Supplementary Data/Random Forest/00. RF Importance Values - Progression - Formatted.xlsx')
```

```{r Plot Importance Values}
# Taxonomy
p1 = imp3 %>% filter(Data=='Covar') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p2 = imp3 %>% filter(Data=='Taxonomy') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Taxonomy') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p3 = imp3 %>% filter(Data=='Taxonomy & Covar') %>% mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=18) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('Taxonomy & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

prog.taxa = ggarrange(p1,p2,p3, ncol=3, widths = c(4,6,8))

# MetaCyc
p4 = imp3 %>% filter(Data=='MetaCyc') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   ylab('Importance') + ggtitle('MetaCyc') + ggeasy::easy_center_title() +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='MetaCyc & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('MetaCyc & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

prog.meta = ggarrange(p4,p5, ncol=2, widths = c(6,8))

# KO
p4 = imp3 %>% filter(Data=='KO') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('KO') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='KO & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('KO & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

prog.ko = ggarrange(p4,p5, nrow=2)

# EC
p4 = imp3 %>% filter(Data=='EC') %>%
  .[-which(duplicated(.$Feature)),] %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('EC') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='EC & Covar') %>%
  .[-which(duplicated(.$Feature)),] %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('EC & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

prog.ec = ggarrange(p4,p5, nrow=2)

# COG
p4 = imp3 %>% filter(Data=='COG') %>% 
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('COG') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))
p5 = imp3 %>% filter(Data=='COG & Covar') %>%
  mutate(Feature=factor(Feature,levels=.$Feature)) %>% 
    ggplot(aes(y = Importance, x = Feature, fill=-Importance)) + geom_col() +
    
    theme_classic(base_size=16) +
    theme(legend.position = 'none')  +  
  scale_y_continuous(expand = expansion(mult = c(0,0.15))) +xlab(NULL) + ylab('Importance') + ggtitle('COG & Covar') + ggeasy::easy_center_title() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_gradientn(colors = gray.colors(100,start = 0.2, end = .6, rev = F))

prog.cog = ggarrange(p4,p5, nrow=2)
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Format for figures

```{r Figure 5cd}
p5 = ggarrange(plotlist = list(plot.taxa, plot.meta, prog.taxa, prog.meta),
               nrow = 4, heights = c(1,1.2,1,1.2))
ggsave('../Results/Figures/Figure 5c,d - RF Importance Scores.pdf', height = 14, width = 12)
```

```{r Figure S6}
pS6a = ggarrange(plotlist = list(plot.cog, plot.ko, plot.ec),
               nrow = 3, heights = c(1,1,1))
pS6b = ggarrange(plotlist = list(prog.cog, prog.ko, prog.ec),
               nrow = 3, heights = c(1,1,1))
pS6 = ggarrange(plotlist = list(pS6a, pS6b), ncol = 2)
ggsave('../Results/Supp Figures/Figure S6 - RF Importance Scores.pdf', height = 18, width = 24)
```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Figure 5efg

```{r}
# Calculate most important features
imp5 = imp4 %>% group_by(Data) %>% 
  summarise(IQR = IQR(Importance),
            q25 = quantile(Importance,0.25),
            q75 = quantile(Importance,0.75)) %>% 
  ungroup() %>% 
  right_join(imp4) %>% 
  mutate(outlier = Importance > q75+(1.5*IQR)) %>% 
  group_by(Data) %>% arrange(-Importance) %>% 
  mutate(biggest = Importance==Importance[1L]) %>% ungroup() %>% 
  mutate(plot = biggest | outlier)

# Main figure
imp5f = imp5 %>% filter(Data == 'Covar' | str_detect(Data,'Taxonomy') | str_detect(Data,"MetaCyc")) %>% 
  filter(plot==T)
# Other datasets
imp5f2 = imp5 %>% filter(!(Data %in% imp5f$Data)) %>% filter(plot==T) 
# if EC, need to change the names so it matches up with other dataset
imp5f2$Feature[str_detect(imp5f2$Data,'EC')] = sapply(imp5f2$Feature[str_detect(imp5f2$Data,'EC')],
                                                      function(x) str_replace_all(x,'[^a-zA-Z\\d\\s:]','.') %>% str_replace_all(': ','..'))

# MAIN FIGURE PLOTS ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Load data from RF
rf.tax = read.csv('../Reference Files/FOR RF - Progression - Taxonomy.csv')
rf.meta = read.csv('../Reference Files/FOR RF - Progression - MetaCyc.csv')

# Plot
temp = rf.tax %>% select(Sample,Estimate, Progression, any_of(imp5f$Feature)) %>% 
  left_join(rf.meta %>% select(Sample, 'PWY66.409..superpathway.of.purine.nucleotide.salvage')) %>% 
  mutate(Progression = factor(.$Progression, levels = c('Slow','Med','Fast')))

p1 = temp %>% 
  ggplot(aes(Progression, s__Faecalibacterium_prausnitzii,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  ylab('F. prausnitzii') +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  scale_fill_grey(start = 0.5, end = 1)

p2 = temp %>% 
  ggplot(aes(Progression,PWY66.409..superpathway.of.purine.nucleotide.salvage,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  ylab('Purine Nucleotide Salvage') +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  scale_fill_grey(start = 0.5, end = 1)

p3 = temp %>% 
  ggplot(aes(Progression,coffee_per_day,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0.2,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  ylab('Cups Coffee/Day') +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  scale_fill_grey(start = 0.5, end = 1)

p5efg = ggarrange(plotlist = list(p3,NULL,p1,NULL,p2),ncol = 1, heights = c(8,2,8,2,8))

ggsave(plot=p5efg,"../Results/Figures/Figure 5efg - RF Boxplots.pdf", height=15,width=5)

# SUPP FIGURE PLOTS ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Load data from RF and fix names
rf.cog = read.csv('../Reference Files/FOR RF - Progression - COG.csv')
names(rf.cog)[8:ncol(rf.cog)] = names(rf.cog)[8:ncol(rf.cog)] %>% sapply(function(x)x %>% str_split('[.][.]') %>% .[[1]] %>% .[1])
rf.ec = read.csv('../Reference Files/FOR RF - Progression - EC.csv')
names(rf.ec)[8:ncol(rf.ec)] = names(rf.ec)[8:ncol(rf.ec)] %>% sapply(function(x)x %>% str_replace_all('[^a-zA-Z\\d\\s:]','.'))
names(rf.ec) = sapply(names(rf.ec), function(x){ ifelse(str_sub(x,end=1)=='X',str_sub(x,start=2),x)})
rf.ko = read.csv('../Reference Files/FOR RF - Progression - KO.csv')
names(rf.ko)[8:ncol(rf.ko)] = names(rf.ko)[8:ncol(rf.ko)] %>% sapply(function(x)x %>% str_split('[.][.]') %>% .[[1]] %>% .[1])

# Plot
temp = rf.cog %>% select(Sample,Estimate, Progression, any_of(imp5f2$fixed)) %>% 
  left_join(rf.ec %>% select(Sample, any_of(imp5f2$Feature))) %>% 
  left_join(rf.ko %>% select(Sample, any_of(imp5f2$fixed))) %>% 
  mutate(Progression = factor(.$Progression, levels = c('Slow','Med','Fast'))) %>% 
  pivot_longer(cols = names(.)[4:ncol(.)], names_to = 'Feature',values_to = 'value')
temp$Feature = sapply(temp$Feature, function(x) ifelse(x %in% imp5f2$Feature,imp5f2$fixed[which(x == imp5f2$Feature)],x)) 
temp = temp %>% 
  left_join(imp5f2 %>% select(Data,fixed) %>% rename(Feature=fixed) %>% 
              mutate(Data = str_remove(Data,' & Covar')) %>% unique)
```

```{r}
p1 = temp %>% filter(Data=='COG') %>% 
  ggplot(aes(Progression, value,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  facet_wrap('Feature',nrow=1) + xlab(NULL) + ylab('') +
  scale_fill_grey(start = 0.5, end = 1)


p2 = temp %>% filter(Data=='EC') %>% 
  ggplot(aes(Progression, value,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  ylab('Feature Abundance') +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  facet_wrap('Feature',nrow=1) + xlab(NULL) +
  scale_fill_grey(start = 0.5, end = 1)


p3 = temp %>% filter(Data=='KO') %>% 
  ggplot(aes(Progression, value,fill = Progression)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0,width=0.2) +
  theme_classic(base_size=18) +
  ggpubr::stat_compare_means(comparisons = list(c('Slow','Med'),c('Med','Fast'),c('Slow','Fast')), size=6) +
  theme(legend.position = 'none') +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  facet_wrap('Feature',nrow=1) + ylab('') + xlab('MDS-UPDRS Part IV Progression') +
  scale_fill_grey(start = 0.5, end = 1)

ggarrange(plotlist=list(p1,p2,p3),nrow=3,heights = c(1,1,1.1))
ggsave("../Results/Supp Figures/Figure S7 - RF Boxplot.pdf", height=10,width=11)
```